---
title: "Regressing emissions from vehicle and road data"
author: "James David Smith"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)

### Loading any missing libraries

library(rvest)
library(stringr)
library(raster)
library(rgdal)
library(rgeos)
library(ggplot2)
library(sf)
library(rmarkdown)
library(rasterVis)
library(knitr)
library(kableExtra)
library(reshape2)
library(mapview)
require(spatialEco)
require(sp)
```

## Import Grid points

Import points which represent 1km grid centroids of a grid covering London

```{r import_grid_points}

grid        <- read.csv('X:/Projects/LAEI2013_TfL_Strategy/z_GridsForJames.csv')
grid        <- grid[,c('GridIdEx', 'Easting', 'Northing')]

names(grid) <- c('grid_exact_cut', 'easting', 'northing')

```

This gives a table of **`r nrow(grid)` rows**, show below.

```{r, echo=F}
kable(head(grid)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "left")
```

## Import Traffic flows

Now import traffic flows by toid and grid exact cut, and edit/clean to get what we need

```{r import_traffic_flows}

traffic_flows                 <- read.csv('X:/Projects/LAEI2013_Gdrive/TrafficFlows/Major_WithAdjustedIbus_AADT/z_AADT_ExactCut_WithIbus_2013.csv')
names(traffic_flows)          <- tolower(names(traffic_flows))

drop                          <- c('year', 'location_exactcut', 'boroughname_exactcut', 'tlrn', 'irr', 'motorwaynumber', 'petrolcar', 'dieselcar', 'electriccar', 'petrollgv', 'diesellgv', 'electriclgv', 'ltbus', 'coach')
traffic_flows                 <- traffic_flows[, !(names(traffic_flows) %in% drop)]

traffic_flows$toid            <- as.character(traffic_flows$toid)

traffic_flows$total_vehicles  <- (traffic_flows$motorcycle + traffic_flows$taxi + traffic_flows$car + traffic_flows$busandcoach + traffic_flows$lgv + traffic_flows$rigid2axle + traffic_flows$rigid3axle + traffic_flows$rigid4axle +     traffic_flows$artic3axle + traffic_flows$artic5axle + traffic_flows$artic6axle)

traffic_flows$total_light     <- (traffic_flows$motorcycle + traffic_flows$taxi + traffic_flows$car)

traffic_flows$total_heavy     <- (traffic_flows$busandcoach + traffic_flows$lgv + traffic_flows$rigid2axle + traffic_flows$rigid3axle + traffic_flows$rigid4axle + traffic_flows$artic3axle + traffic_flows$artic5axle + traffic_flows$artic6axle)

drop                          <- c('motorcycle', 'taxi', 'car', 'busandcoach', 'lgv', 'rigid2axle', 'rigid3axle', 'rigid4axle', 'artic3axle', 'artic5axle', 'artic6axle')
traffic_flows                 <- traffic_flows[, !(names(traffic_flows) %in% drop)]

```

This gives a table of **`r nrow(traffic_flows)` rows**, show below.

```{r, echo=F}
kable(head(traffic_flows)) %>%  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "left")
```

## Import Emissions

Import emissions by toid and grid exact cut

```{r import_emissions}

emissions                     <- read.csv('X:/Projects/LAEI2013_TfL_Strategy/FilesForHerald4/EmissionsByLink/INPUT_RESULTS_LAEI_ExactCutIntersect_Major_2013_LAEI_V117.csv')
names(emissions)              <- tolower(names(emissions))

drop                          <- c('emissions', 'year', 'petrolcar', 'dieselcar', 'petrollgv', 'diesellgv', 'ltbus', 'coach', 'electriccar', 'electriclgv')
emissions                     <- emissions[, !(names(emissions) %in% drop)]

emissions$total_emissions     <- emissions$motorcycle + emissions$taxi + emissions$car + emissions$busandcoach + emissions$lgv + emissions$rigid + emissions$artic + emissions$rigid2axle + emissions$rigid3axle + emissions$rigid4axle + emissions$artic3axle + emissions$artic5axle + emissions$artic6axle

emissions$toid                <- as.character(emissions$toid)

drop                          <- c('motorcycle', 'taxi', 'car', 'busandcoach', 'lgv', 'rigid', 'artic', 'rigid2axle', 'rigid3axle', 'rigid4axle', 'artic3axle', 'artic5axle', 'artic6axle')
emissions                     <- emissions[, !(names(emissions) %in% drop)]

emissions                     <- emissions[!(emissions$pollutant %in% c('TB_PM10_SW', 'TB_PM25_SW')),]

```

Not going to need all the constiuent parts of the emissions, just the total by each. So aggregate up.

```{r aggregate_emissions}

emissions$pollutant           <- as.character(emissions$pollutant)

emissions[grep('pm25', tolower(emissions$pollutant)),'pollutant'] <- 'pm25'
emissions[grep('pm10', tolower(emissions$pollutant)),'pollutant'] <- 'pm10'

emissions$gridid            <- as.character(emissions$gridid)
emissions$grid_exactcut_id  <- as.character(emissions$grid_exactcut_id)
emissions$dotref            <- as.character(emissions$dotref)

emissions$pollutant         <- tolower(emissions$pollutant)

emissions <- aggregate(total_emissions ~ gridid + toid + grid_exactcut_id + dotref + length + pollutant, data=emissions, FUN=sum)


```

This gives a table of **`r nrow(emissions)` rows**, show below.

```{r, echo=F}
kable(head(emissions)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "left")
```

## Link traffic flows to grids

```{r joining_traffic_to_grid}

grid_traffic              <- merge(traffic_flows, grid, by.x = c('grid_exactcut_id'), by.y = c('grid_exact_cut'))
rm(traffic_flows)
grid_traffic              <- grid_traffic[order(grid_traffic$toid),]
```

Looks like this

```{r, echo=F}
kable(head(grid_traffic)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "left")
```

Now take the mean traffic flow, on each toid, by grid square (using easting and northing)

```{r mean_traffic_by_toid_grid}
grid_traffic <- aggregate(data = grid_traffic, FUN=mean, cbind(speed, total_vehicles, total_light, total_heavy)~ easting+northing+toid)
```

Now looks like this

```{r, echo=F}
kable(head(grid_traffic[order(grid_traffic$toid),])) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "left")
```

## Link emissions to grids

```{r joining_grids_to_emissions}

emissions_grids              <- merge(emissions, grid, by.x = 'grid_exactcut_id', by.y = 'grid_exact_cut')
rm(emissions)
emissions_grids              <- emissions_grids[order(emissions_grids$toid),]

```

## Link emissions, to traffic, to grids

```{r joining_grids_to_emissions_to_traffic}

all_data                <- merge(emissions_grids, grid_traffic, by.x = c('toid', 'easting', 'northing'), by.y = c('toid', 'easting', 'northing'))

rm(grid, grid_traffic, emissions_grids,drop)

```

Example of final data for NO2 and one grid

```{r, echo=F}
kable(all_data[all_data$easting == 512500 & all_data$northing == 198500 & all_data$pollutant == 'no2',]) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position = "left")
```

## Aggregate 1km grids to 10km

First sum emissions by the 1km, to 1km grid, and create raster stacks from them

```{r sum_up_emissions_by_1km}
ukgrid = "+init=epsg:27700"

one_km_emissions <- aggregate(data = all_data, total_emissions ~ pollutant + easting + northing, FUN=sum)

sources          <- as.character((unique(one_km_emissions$pollutant)))

for (i in 1:length(sources)) {
  
  temp              <- one_km_emissions[one_km_emissions$pollutant == sources[i],]
  coordinates(temp) <- ~ easting + northing
  temp$pollutant    <- NULL
  gridded(temp)     <- TRUE
  proj4string(temp) <- CRS(ukgrid)
  temp              <- raster(temp)
  names(temp)       <- tolower(sources[i])

  if (i == 1) { one_km_emissions_raster <- temp} else {one_km_emissions_raster <- stack(one_km_emissions_raster, temp)}
  
}

```

**NO2 1km plot**

```{r, echo=F}
plot(one_km_emissions_raster$no2)
```

Now make a 10km plot

```{r make_ten_km_grid}
ten_km_emissions_raster  <- aggregate(one_km_emissions_raster, fact = 10, fun=sum)

rm(one_km_emissions, one_km_emissions_raster, temp, i)
```

**NO2 10km plot**

```{r, echo=F}
plot(ten_km_emissions_raster$no2)
```

Now need to think about how these concentrations at 10km level, relate to the road link that contributed them, and get the relationship between them. First convert the raster to polygons.

```{r linking_10km_back_to_roads}
ten_km_emissions_polygons <- rasterToPolygons(ten_km_emissions_raster)
```

**NO2 10km polygons plot**

```{r, echo=F}
mapview(ten_km_emissions_polygons, map.types = 'OpenStreetMap')
```

## Link TOIDS to 10km grids
Turn the originl data into points, so that can use these to extract the values from the 10km polygons
```{r map_originals_to_points}

coordinates(all_data)           <- ~ easting + northing
proj4string(all_data)           <- CRS(ukgrid)

```

## Calculate each TOIDS contribution 

This code now looks at each road, sees which overall 10km grid it is in, and then divides each pollutant by the total of that pollutant in the square. Puts the result into 'ten_km_contribution' variable.

```{r get_point_polygon_and_divide}

result                                                  <- point.in.poly(all_data, ten_km_emissions_polygons)

result                                                  <- data.frame(result)

result$ten_km_contribution                              <- NA

result[result$pollutant == 'no2',]$ten_km_contribution    <- result[result$pollutant == 'no2',]$total_emissions    / result[result$pollutant == 'no2',]$no2
result[result$pollutant == 'nox',]$ten_km_contribution    <- result[result$pollutant == 'nox',]$total_emissions    / result[result$pollutant == 'nox',]$nox
result[result$pollutant == 'pm10',]$ten_km_contribution   <- result[result$pollutant == 'pm10',]$total_emissions   / result[result$pollutant == 'pm10',]$pm10
result[result$pollutant == 'pm25',]$ten_km_contribution   <- result[result$pollutant == 'pm25',]$total_emissions   / result[result$pollutant == 'pm25',]$pm25

rm(all_data)


```

Some toid exact cuts (`r nrow(result[result$total_emissions > 0,])`) have zero emissions, remove those from the total of `r nrow(result)`.

```{r remove_zero_roads}

result <- result[result$total_emissions > 0,]

```

```{r aggregate_to_dotref}

result <- merge(aggregate(data = result, cbind(length, total_emissions, ten_km_contribution) ~ dotref + pollutant, sum),
          aggregate(data = result, cbind(speed, total_vehicles, total_light, total_heavy) ~ dotref + pollutant, mean),
          by = c('dotref', 'pollutant'))

```

## Exploring data relationships

### Histogram of 10km emission contributions

```{r emissions_histogram, fig.width=12, }

ggplot(result, aes(ten_km_contribution)) +
  geom_histogram(bins = 60)

```

### Total vehicles v. emission contribution
```{r ggplot_of_variables1, fig.width=10, echo=F}

ggplot(result, aes(total_vehicles, ten_km_contribution)) +
  geom_point(alpha = 0.2, colour = 'blue', fill = 'blue', size=0.5) +
  facet_wrap(.~pollutant, scales = 'free') +
  geom_smooth(colour = 'red', method = 'lm') +
  xlab('Number of vehicles (all)') +
  ylab('Contribution to 10km grid emissions')

```

### Light vehicles v. emission contribution
```{r ggplot_of_variables2, fig.width=12, echo=F}

ggplot(result, aes(total_light, ten_km_contribution)) +
  geom_point(alpha = 0.2, colour = 'blue', fill = 'blue', size=0.5) +
  facet_wrap(.~pollutant, scales = 'free') +
  geom_smooth(colour = 'red', method = 'lm') +
  xlab('Number of light vehicles') +
  ylab('Contribution to 10km grid emissions')

```

### Heavy vehicles v. emission contribution
```{r ggplot_of_variables3, fig.width=12, echo=F}

ggplot(result, aes(total_heavy, ten_km_contribution)) +
  geom_point(alpha = 0.2, colour = 'blue', fill = 'blue', size=0.5) +
  facet_wrap(.~pollutant, scales = 'free') +
  geom_smooth(colour = 'red', method = 'lm') +
  xlab('Number of heavy vehicles') +
  ylab('Contribution to 10km grid emissions')

```

### Speed v. emission contribution
```{r ggplot_of_variables4, fig.width=12, echo=F}

ggplot(result, aes(speed, ten_km_contribution)) +
  geom_point(alpha = 0.2, colour = 'blue', fill = 'blue', size=0.5) +
  facet_wrap(.~pollutant, scales = 'free') +
  geom_smooth(colour = 'red', method = 'lm') +
  xlab('Speed of vehicles') +
  ylab('Contribution to 10km grid emissions')

```

### Length v. emission contribution**
```{r ggplot_of_variables5, fig.width=12, echo=F}

ggplot(result, aes(length, ten_km_contribution)) +
  geom_point(alpha = 0.2, colour = 'blue', fill = 'blue', size=0.5) +
  facet_wrap(.~pollutant, scales = 'free') +
  geom_smooth(colour = 'red', method = 'lm') +
  xlab('Length of toid') +
  ylab('Contribution to 10km grid emissions')

```

## Create training and testing data

Take 80% of the data as a training data set

```{r make_training_model}

training_index <- sample(nrow(result), round(nrow(result) * 0.8,0))

training_data <- result[training_index,]

```

Seperate the other 20% as a testing data set

```{r make_testing_data}

testing_data <- result[-training_index,]

```

## Make linear regression model

Make a liner regression model using the training data

```{r make_regressions}

model_results <- data.frame(pollutant = NA,
                            intercept = as.numeric(''),
                            length_coef = as.numeric(''),
                            speed_coef = as.numeric(''),
                            total_light_coef = as.numeric(''),
                            total_heavy_coef = as.numeric(''),
                            stringsAsFactors = F)

model_results_list <- list()

for (i in 1:length(sources)) {
  
  model <- lm(ten_km_contribution ~ length + speed + total_light + total_heavy,
             data=training_data[training_data$pollutant == sources[i],])

  model_results_list[[i]] <- model
  
  model_results[i,] <- c(sources[i], as.numeric(coef(model)[1]), as.numeric(coef(model)[2]), as.numeric(coef(model)[3]), as.numeric(coef(model)[4]), as.numeric(coef(model)[5]))
  
  rm(model)
}

model_results$intercept         <- as.numeric(model_results$intercept)
model_results$length_coef       <- as.numeric(model_results$length_coef)
model_results$speed_coef        <- as.numeric(model_results$speed_coef)
model_results$total_light_coef  <- as.numeric(model_results$total_light_coef)
model_results$total_heavy_coef  <- as.numeric(model_results$total_heavy_coef)

```

## Model result
* `r toupper(sources[1])` adjusted R-squared: `r summary(model_results_list[[1]])$adj.r.squared`
* `r toupper(sources[2])` adjusted R-squared: `r summary(model_results_list[[2]])$adj.r.squared`
* `r toupper(sources[3])` adjusted R-squared: `r summary(model_results_list[[3]])$adj.r.squared` 
* `r toupper(sources[4])` adjusted R-squared: `r summary(model_results_list[[4]])$adj.r.squared`

## Apply linear regression model
Now use the regression equation on the testing 20% data

```{r test_regression}

testing_data$prediction <- NA

for (i in 1:length(sources)) {
  testing_data[testing_data$pollutant == sources[i],'prediction'] <-    model_results[model_results$pollutant == sources[i],]$intercept+(
                                                                  (testing_data[testing_data$pollutant == sources[i],]$length      * model_results[model_results$pollutant == sources[i],]$length_coef) +
                                                                  (testing_data[testing_data$pollutant == sources[i],]$speed       * model_results[model_results$pollutant == sources[i],]$speed_coef) +
                                                                  (testing_data[testing_data$pollutant == sources[i],]$total_light * model_results[model_results$pollutant == sources[i],]$total_light_coef) +
                                                                  (testing_data[testing_data$pollutant == sources[i],]$total_heavy * model_results[model_results$pollutant == sources[i],]$total_heavy_coef))
}

```

## Results 
Plot the prediction against the actuals for the testing 20%

```{r plot_predictions, fig.width=12, fig.height = 40, echo=F}

ggplot(testing_data, aes(ten_km_contribution, prediction)) +
  geom_point() +
  facet_wrap(.~pollutant, scales = 'free', ncol = 1)
```
